Commutative character sheaves and geometric types for supercuspidal representations

### 2019-07-02 Geometric types.tex
Introduction and Henselian

\begin{abstract}
We show that some types for supercuspidal representations of tamely ramified $p$-adic groups that appear in Jiu-Kang Yu's work are geometrizable.
To do this we must find the function-sheaf dictionary for one-dimen\-sion\-al characters of arbitrary smooth group schemes over finite fields.  
In previous work we considered the case of commutative smooth group schemes and found that the standard definition of character sheaves produced a dictionary with a nontrivial kernel.  
In this paper we give a modification of the category of character sheaves that remedies this defect, and is also extensible to non-commutative groups.  
We then use these \emph{commutative character sheaves} to geometrize the linear characters that appear in the types introduced by Jiu-Kang Yu, assuming that the character vanishes on a certain derived subgroup.
We combine these sheaves with Lusztig's character sheaves on reductive algebraic groups over finite fields and the geometrization of the Weil representation found by Gurevich and Hadani, to define \emph{geometric types} for supercuspidal representations of tamely ramified $p$-adic groups.
\end{abstract}

@@

As proved by Ju-Lee Kim in \cite{kim:07a}, all irreducible supercuspidal representations of tamely ramified $p$-adic groups for sufficiently large $p$ can be built from ``data'' introduced by Jiu-Kang Yu in \cite{yu:01a}*{\S 15}.
While the type, in the sense of Bushnell \& Kutzko \cite{bushnell-kutzko:98a}, of a supercuspidal representation built from Yu data can be constructed directly from the datum, it is convenient to consider an intermediate object, introduced in \cite{yu:01a}*{Remark 15.4}, which we call a \emph{Yu type datum}. 
Yu type data are studied in \cite{Yu:models}, which concludes with the following observation.
\begin{quotation}
{\it Therefore, up to some linear characters, all the ingredient representations 
 are on groups of the form $\underline{H}(\mathcal{O})$, where $\underline{H}$ is a smooth group scheme over [a henselian discrete valuation ring with finite residue field $\kappa$] $\mathcal{O}$, and the representations are inflated from $\underline{H}(\kappa)$. These results suggest that algebraic geometry and group schemes should play an
important role in the representation theory of $p$-adic groups.}
\end{quotation}

@@

\subsection{Quasicharacters of smooth group schemes over certain henselian traits}\label{ssec:quasicharacters}

>>>>

\begin{abstract}
We show that some types for supercuspidal representations of tamely ramified $p$-adic groups that appear in Jiu-Kang Yu's work are geometrizable.
To do this we must find the function-sheaf dictionary for one-dimen\-sion\-al characters of arbitrary smooth group schemes over finite fields.  
In previous work we considered the case of commutative smooth group schemes and found that the standard definition of character sheaves produced a dictionary with a nontrivial kernel.  
In this paper we give a modification of the category of character sheaves that remedies this defect, and is also extensible to non-commutative groups.  
We then use these \emph{commutative character sheaves} to geometrize the linear characters that appear in the types introduced by Jiu-Kang Yu, assuming that the character vanishes on a certain derived subgroup.
To define \emph{geometric types} we combine commutative character sheaves with Gurevich and Hadani's geometrization of the Weil representation and Lusztig's character sheaves.
\end{abstract}

@@

As proved by Ju-Lee Kim in \cite{kim:07a}, all irreducible supercuspidal representations of tamely ramified $p$-adic groups for sufficiently large $p$ can be built from ``data'' introduced by Jiu-Kang Yu in \cite{yu:01a}*{\S 15}.
While the type, in the sense of Bushnell \& Kutzko \cite{bushnell-kutzko:98a}, of a supercuspidal representation built from Yu data can be constructed directly from the datum, it is convenient to consider an intermediate object, introduced in \cite{yu:01a}*{Remark 15.4}, which we call a \emph{Yu type datum}. 
Yu type data are studied in \cite{Yu:models}, which concludes with the following observation.
\begin{quotation}
{\it Therefore, up to some linear characters, all the ingredient representations 
 are on groups of the form $\underline{H}(\mathcal{O})$, where $\underline{H}$ is a smooth group scheme over [a Henselian discrete valuation ring with finite residue field $\kappa$] $\mathcal{O}$, and the representations are inflated from $\underline{H}(\kappa)$. These results suggest that algebraic geometry and group schemes should play an
important role in the representation theory of $p$-adic groups.}
\end{quotation}

@@

\subsection{Quasicharacters of smooth group schemes over certain Henselian traits}\label{ssec:quasicharacters}
